Evolution of the AGN UV luminosity function from redshift 7.5

### 2017-09-21 Paper/paper.tex
Update text

The 1~Ry emissivity can now be used to calculate the ionizing flux and
photoionization rate contributed by quasars.  The frequency dependence
above 1~Ry is given by Equation~(\ref{eqn:sed}) or
(\ref{eqn:sed_faint}), so that for $\nu > \nu_{912}$
\begin{equation}
  \epsilon_\nu = \epsilon_{912}\left(\frac{\nu}{\nu_{912}}\right)^\alpha,
  \label{eqn:epsilon_freq}
\end{equation}
where $\alpha=-1.70$ (Equation~\ref{eqn:sed}) or $-0.56$ (Equation~\ref{eqn:sed_faint}).

The emissivity calculated in Equation~(\ref{eqn:epsilon}) is a source
property.  We are now interested in what flux this emissivity produces
in the surrounding intergalactic medium.  This brings in the medium's
properties also, via the radiative transfer equation.  The flux is
\citep{2012ApJ...746..125H}
\begin{multline}
  j(\nu_0, z_0)=\frac{1}{4\pi}\int_{z_0}^\infty dz\frac{dl}{dz}\frac{(1+z_0)^3}{(1+z)^3}\epsilon(\nu,z)\\
  \times\exp{(-\tau_\mathrm{eff}(\nu_0, z_0, z))},
  \label{eqn:flux}
\end{multline}
where
\begin{equation}
  \nu = \nu_0\left(\frac{1+z}{1+z_0}\right).
\end{equation}
We estimate the effective optical depth as
\begin{equation}
  \tau_\mathrm{eff}(\nu_0, z_0, z) = \int_{z_0}^z dz^\prime\int_0^\infty dN_\mathrm{HI} f(N_\mathrm{HI}, z^\prime) (1-e^{-\tau_\nu}),
\end{equation}
where $\tau_\nu=\sigma_\nu N_\mathrm{HI}$ and the column density
distribution $f(N_\mathrm{HI}, z)$ is taken to be a power law \citep{2013MNRAS.436.1023B}
\begin{equation}
  f(N_\mathrm{HI}, z) = \frac{A}{N_\mathrm{LL}}\left(\frac{N_\mathrm{HI}}{N_\mathrm{LL}}\right)^{-\beta_N}\left(\frac{1+z}{4.5}\right)^{-\beta_z},
\end{equation}
with $A=0.93$, $\beta_N=1.33$, $\beta_z=1.92$, and
$N_\mathrm{LL}=10^{17.2}$~cm$^{-2}$.  The hydrogen photoionization
rate is then given by
\begin{equation}
  \Gamma_\mathrm{HI}=4\pi\int_{\nu_{912}}^\infty d\nu \frac{j(\nu,z)}{h\nu} \sigma(\nu),
\end{equation}

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{g_18.pdf}
    \includegraphics[width=0.5\textwidth]{g_20.pdf}
    \end{tabular}
  \end{center}
  \caption{AGN contribution to the hydrogen photoionisation rate,
    assuming 100\% escape fraction.  Model luminosity functions are
    integrated down to $M_{1450}=-18$ in the left panel and $-20$ in
    the right panel.}
  \label{fig:gammapi}
\end{figure*}

\subsection{Local source approximation}

In the local source approximation, Equation~(\ref{eqn:flux}) becomes
\begin{equation}
  j(\nu_0, z_0) = \frac{1}{4\pi}\lambda_\mathrm{mfp}(\nu_0, z_0)\epsilon(\nu_0, z_0),
\end{equation}
where $\lambda_\mathrm{mfp}$ is the mean free path.  The hydrogen
photoionization rate is given by
\begin{equation}
  \Gamma_\mathrm{HI}=4\pi\int_{\nu_{912}}^\infty d\nu \frac{j_\nu}{h\nu} \sigma(\nu),
  \label{eqn:gammapi}
\end{equation}
where $\sigma$ is the ionization cross-section
\begin{equation}
  \sigma(\nu) = \sigma_0\left(\frac{\nu}{\nu_{912}}\right)^{-3},
  \label{eqn:sigma}
\end{equation}
with $\sigma_0=6.3\times 10^{-18}$~cm$^2$ \citep{2006agna.book.....O}.
In the redshift range $z=2.3$--$5.5$, the mean free path is measured
to be \citep{2014MNRAS.445.1745W}
\begin{equation}
  \lambda_\mathrm{mfp}(\nu, z)= \lambda_0\left(\frac{1+z}{5}\right)^{-5.4}\left(\frac{\nu}{\nu_{912}}\right)^{1.5},
  \label{eqn:mfp}
\end{equation}
where the frequency dependence comes from the assumed column density
distribution $f(N_\mathrm{HI})\propto N_\mathrm{HI}^{-1.5}$.  We
extrapolate the redshift dependence in Equation~(\ref{eqn:mfp}) at
$z<2.3$ and $z>5.5$.  Combining Equations~(\ref{eqn:epsilon_freq}),
(\ref{eqn:sigma}), and (\ref{eqn:mfp}), Equation~(\ref{eqn:gammapi})
gives
\begin{multline}
  \Gamma_\mathrm{HI}=4.6\times 10^{-13} \mathrm{s}^{-1} \left(\frac{\epsilon_{912}}{10^{24}\mathrm{erg s^{-1} Hz^{-1} cMpc^{-3}}}\right)\\
  \times\left(\frac{1}{1.5-\alpha}\right)\left(\frac{1+z}{5}\right)^{-2.4}.
\end{multline}
Note that this equation refers to the \emph{comoving} emissivity.  The
local source approximation works well for $z>3$ \citep{2013MNRAS.436.1023B}.

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth,keepaspectratio]{boss.pdf}
%%   \end{center}
%%   \caption{Comparison of LF derived by our method with the published
%%     LF of \citet{2013ApJ...773...14R} for the BOSS DR9 color-selected
%%     sample. There are 23,301 quasars in this sample ($2.2<z<3.5$).  Of
%%     these, 231 quasars have $V_i=0$ and are dropped from the
%%     analysis.}
%%   \label{fig:boss}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth,keepaspectratio]{sdss.pdf}
%%   \end{center}
%%   \caption{Comparison with the published LF of
%%     \citet{2006AJ....131.2766R} for SDSS DR3.  There are 15,343
%%     quasars in this sample, of which 155 quasars have $z<0.3$ and 9
%%     quasars have $z>5$.  Of the remaining 15179 quasars with
%%     $0.3<z<5$, 60 quasars have $V_i=0$ and are dropped from the
%%     analysis.  As a result, the green points above contain 15,119
%%     quasars.  The red circles contain 14,953 quasars.}
%%   \label{fig:sdss}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=0.6\textwidth,keepaspectratio]{mcgreer.pdf}
%%   \end{center}
%%   \caption{Comparison with the published LF of
%%     \citet{2013ApJ...768..105M} for $4.7\leq z<5.1$.  There are 148
%%     quasars in their DR7 sample and 52 quasars in the Stripe 82
%%     sample. Of these 1 quasar from the Stripe 82 sample has $V_i=0$.
%%     The published samples of \citet{2013ApJ...768..105M} have 146 and
%%     51 quasars.}
%%   \label{fig:mcgreer}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=0.6\textwidth,keepaspectratio]{glikman.pdf}
%%   \end{center}
%%   \caption{Comparison with the published LF of
%%     \citet{2011ApJ...728L..26G} for $3.5<z<5.2$.  There are 24 
%%     quasars in their two samples (NDWFS and DLS).}
%%   \label{fig:mcgreer}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth]{glikman3.pdf}
%%   \end{center}
%%   \caption{Attempts at reproducing Glikman's published LF.  Panels (a)
%%     and (b) use her volume calculations.  Panels (c)--(e) use our
%%     volume calculations.  Various corrections incorporated by Gabor
%%     are in panels (c)--(e).  Red open circles in all panels are
%%     Glikman's published values; I cannot reproduce them in any of the
%%     panels.  All luminosity function in this version of the draft use
%%     Gabor's October data from panel (d).}
%%   \label{fig:glikman}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth]{giallongo2.pdf}
%%   \end{center}
%%   \caption{Giallongo's LF when selection function is taken to be
%%     $\phi_\mathrm{corr}/\phi$ from their paper.  This plots shows that
%%     additional correction is required, which has been incorporated in
%%     our LFs.}
%%   \label{fig:giallongo}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth,keepaspectratio]{mosaic2.pdf}
%%   \end{center}
%%   \caption{Individual luminosity function fits for $z=2.6$ to $6.5$.}
%%   \label{fig:mosaic2}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=0.6\textwidth,keepaspectratio]{rhoqso_allz.pdf}
%%   \end{center}
%%   \caption{AGN number density evolution, showing the systematic offset in BOSS quasars.}
%%   \label{fig:rhoqso_allz}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=0.6\textwidth,keepaspectratio]{rhoqso.pdf}
%%   \end{center}
%%   \caption{AGN number density evolution with some BOSS quasars
%%     removed.}
%%   \label{fig:rhoqso}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth]{evolution_seldata.pdf}
%%   \end{center}
%%   \caption{Parameter evolution with some BOSS quasars removed.  We (1)
%%     use \emph{only} BOSS quasars for $z=2.2$--$3.5$, and (2) use only
%%     $p>0.9$ quasars in the $z>3.7$ sample of Richards.}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth,keepaspectratio]{mosaic_withGlobal.pdf}
%%   \end{center}
%%   \caption{Individual and composite luminosity function fits for
%%     $z=0.3$ to $2.6$.  They do not agree for $2.2 < z < 2.6$ because
%%     composite fit does not consider quasars at those redshifts.}
%%   \label{fig:mosaic_withGlobal}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth,keepaspectratio]{mosaic2_withGlobal.pdf}
%%   \end{center}
%%   \caption{Individual and composite luminosity function fits for
%%     $z=2.6$ to $6.5$.  They do not agree for $2.6 < z < 2.8$ because
%%     composite fit does not consider quasars at those redshifts.}
%%   \label{fig:mosaic2_withGlobal}
%% \end{figure*}

%% \begin{figure*}
%%   \begin{center}
%%     \begin{tabular}{cc}
%%     \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
%%     \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
%%     \end{tabular}
%%   \end{center}
%%   \caption{LyC emissivity of AGN assuming 100\% escape fraction.
%%     Model luminosity functions are integrated down to $M_{1450}=-18$
%%     in the left panel and $-20$ in the right panel.}
%%   \label{fig:gammapi}
%% \end{figure*}

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.7\textwidth,keepaspectratio]{q.pdf}
  \end{center}
  \caption{Helium reionization.}
\end{figure*}

>>>>

The 1~Ry emissivity can now be used to calculate the ionizing flux and
photoionization rate contributed by quasars.  The frequency dependence
above 1~Ry is given by Equation~(\ref{eqn:sed}) or
(\ref{eqn:sed_faint}), so that for $\nu > \nu_{912}$
\begin{equation}
  \epsilon_\nu = \epsilon_{912}\left(\frac{\nu}{\nu_{912}}\right)^\alpha,
  \label{eqn:epsilon_freq}
\end{equation}
where $\alpha=-1.70$ (Equation~\ref{eqn:sed}) or $-0.56$
(Equation~\ref{eqn:sed_faint}).  The corresponding flux in the IGM is
written as a solution of the cosmological radiative transfer equation
as \citep{2012ApJ...746..125H}
\begin{multline}
  j(\nu_0, z_0)=\frac{1}{4\pi}\int_{z_0}^\infty dz\frac{dl}{dz}\frac{(1+z_0)^3}{(1+z)^3}\epsilon(\nu,z)\\
  \times\exp{(-\tau_\mathrm{eff}(\nu_0, z_0, z))},
  \label{eqn:flux}
\end{multline}
where
\begin{equation}
  \nu = \nu_0\left(\frac{1+z}{1+z_0}\right).
\end{equation}
Here the effective optical depth is estimated as
\begin{equation}
  \tau_\mathrm{eff}(\nu_0, z_0, z) = \int_{z_0}^z dz^\prime\int_0^\infty dN_\mathrm{HI} f(N_\mathrm{HI}, z^\prime) (1-e^{-\tau_\nu}),
\end{equation}
where $\tau_\nu=\sigma_\nu N_\mathrm{HI}$ and the column density
distribution $f(N_\mathrm{HI}, z)$ is taken to be a power law \citep{2013MNRAS.436.1023B}
\begin{equation}
  f(N_\mathrm{HI}, z) = \frac{A}{N_\mathrm{LL}}\left(\frac{N_\mathrm{HI}}{N_\mathrm{LL}}\right)^{-\beta_N}\left(\frac{1+z}{4.5}\right)^{-\beta_z},
\end{equation}
with $A=0.93$, $\beta_N=1.33$, $\beta_z=1.92$, and
$N_\mathrm{LL}=10^{17.2}$~cm$^{-2}$.  The hydrogen photoionization
rate is then given by
\begin{equation}
  \Gamma_\mathrm{HI}=4\pi\int_{\nu_{912}}^\infty d\nu \frac{j(\nu,z)}{h\nu} \sigma(\nu),
\end{equation}
where $\sigma$ is the photoionization cross-section.  The result is
shown in Figure~\ref{fig:gammapi}, in which as before the solid black
curve shows the hydrogen photoionisate rate in our best-fit global
model, and the yellow curves show a random sample from the global
posterior.  Also, as before, the right panel of
Figure~\ref{fig:gammapi} shows the result when we only account for
quasars brighter than $M_{1450}=-20$, while the left panel shows the
result when the integration limit is $M_{1450}=-18$.  The red points
with error bars in this figure show the measurements of
\citet{2013MNRAS.436.1023B}, derived from the Lyman-$\alpha$ forest,
and the orange points show the measurements by
\citet{2011MNRAS.412.2543C} from quasar proximity zones.  The dashed
green curve shows the hydrogen photoionisation rate evolution in the
model of \citet{2012ApJ...746..125H}, while the solid green curve
shows the contribution of quasars to it.

The black open circles show the photoionisation rate values derived
from our fits in the individual redshift bins.  However, note that
unlike in the global model it is not possible to do the integral in
Equation~(\ref{eqn:flux}) for these individual fits.  We can only
derive the photoionisatoin rate by assuming what is known as the local
source approximation, that is by assuming that the photons emitted at
redshift $z$ are absorbed at the same redshift without propagation to
large distances.

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{g_18.pdf}
    \includegraphics[width=0.5\textwidth]{g_20.pdf}
    \end{tabular}
  \end{center}
  \caption{AGN contribution to the hydrogen photoionisation rate,
    assuming 100\% escape fraction.  Model luminosity functions are
    integrated down to $M_{1450}=-18$ in the left panel and $-20$ in
    the right panel.  The solid black curve shows the hydrogen
    photoionisate rate in our best-fit global model, and the yellow
    curves show a random sample from the global posterior.  The red
    points with error bars in this figure show the measurements of
    \citet{2013MNRAS.436.1023B}, derived from the Lyman-$\alpha$
    forest, and the orange points show the measurements by
    \citet{2011MNRAS.412.2543C} from quasar proximity zones.  The
    dashed green curve shows the hydrogen photoionisation rate
    evolution in the model of \citet{2012ApJ...746..125H}, while the
    solid green curve shows the contribution of quasars to it. }
  \label{fig:gammapi}
\end{figure*}

In the local source approximation, Equation~(\ref{eqn:flux}) becomes
\begin{equation}
  j(\nu_0, z_0) = \frac{1}{4\pi}\lambda_\mathrm{mfp}(\nu_0, z_0)\epsilon(\nu_0, z_0),
\end{equation}
where $\lambda_\mathrm{mfp}$ is the mean free path.  The hydrogen
photoionization rate is given by
\begin{equation}
  \Gamma_\mathrm{HI}=4\pi\int_{\nu_{912}}^\infty d\nu \frac{j_\nu}{h\nu} \sigma(\nu),
  \label{eqn:gammapi}
\end{equation}
where $\sigma$ is the ionization cross-section
\begin{equation}
  \sigma(\nu) = \sigma_0\left(\frac{\nu}{\nu_{912}}\right)^{-3},
  \label{eqn:sigma}
\end{equation}
with $\sigma_0=6.3\times 10^{-18}$~cm$^2$ \citep{2006agna.book.....O}.
In the redshift range $z=2.3$--$5.5$, the mean free path is measured
to be \citep{2014MNRAS.445.1745W}
\begin{equation}
  \lambda_\mathrm{mfp}(\nu, z)= \lambda_0\left(\frac{1+z}{5}\right)^{-5.4}\left(\frac{\nu}{\nu_{912}}\right)^{1.5},
  \label{eqn:mfp}
\end{equation}
where the frequency dependence comes from the assumed column density
distribution $f(N_\mathrm{HI})\propto N_\mathrm{HI}^{-1.5}$.  We
extrapolate the redshift dependence in Equation~(\ref{eqn:mfp}) at
$z<2.3$ and $z>5.5$.  Combining Equations~(\ref{eqn:epsilon_freq}),
(\ref{eqn:sigma}), and (\ref{eqn:mfp}), Equation~(\ref{eqn:gammapi})
gives
\begin{multline}
  \Gamma_\mathrm{HI}=4.6\times 10^{-13} \mathrm{s}^{-1} \left(\frac{\epsilon_{912}}{10^{24}\mathrm{erg s^{-1} Hz^{-1} cMpc^{-3}}}\right)\\
  \times\left(\frac{1}{1.5-\alpha}\right)\left(\frac{1+z}{5}\right)^{-2.4}.
\end{multline}
Note that this equation refers to the \emph{comoving} emissivity.  The
local source approximation works well for $z>3$, as can be seen in
Figure~\ref{fig:gammapi}, where the points from the individual fits
are in much better agreement with the curves from the global model.
At lower redshift, the photoionisation rate in the individual models
is much higher than that in the global model, because the individual
models overestimate the flux at these redshifts.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth,keepaspectratio]{q.pdf}
  \end{center}
  \caption{Helium reionization.}
\end{figure}
